---
title: "Exploring Palmer Penguins with R"
author: "Vaibhav Sharma"
date: "2026-09-23"
format: html
---

## Introduction

In this post, I explore the Palmer Penguins dataset using R. I compare body mass across the three penguin species and examine differences in their distributions.

Data: [Palmer Penguins](https://allisonhorst.github.io/palmerpenguins/), Palmer Station Antarctica LTER.

## Loading and exploring the data

First, I load the packages and examine the Palmer Penguins dataset.

```{r}
#| message: false
#| warning: false

library(tidyverse)
library(palmerpenguins)

head(penguins)
```

## Missing values

Before analyzing body mass, I check the number of missing values in each variable.

```{r}
colSums(is.na(penguins))
```
## Body mass by species

Next, I calculate the average body mass for each penguin species.

```{r}
body_mass_summary <- penguins |>
  group_by(species) |>
  summarise(
    average_body_mass = mean(body_mass_g, na.rm = TRUE),
    .groups = "drop"
  )

body_mass_summary
```

The summary shows differences in average body mass among the three species. Gentoo penguins have a higher average body mass than Adelie and Chinstrap penguins.
## Distribution of body mass

The following boxplot compares the distribution of body mass across penguin species.

```{r}
#| label: fig-body-mass
#| fig-cap: "Distribution of body mass across Palmer Penguin species."

ggplot(penguins, aes(x = species, y = body_mass_g)) +
  geom_boxplot() +
  labs(
    x = "Species",
    y = "Body Mass (g)",
    title = "Body Mass by Penguin Species"
  ) +
  theme_minimal()
```

The boxplot shows that Gentoo penguins generally have greater body mass than Adelie and Chinstrap penguins. The plot also shows variation in body mass within each species.
